### 2020-06-04 Reportes/Reporte20200601.tex
reporte avance

\documentclass{article}

%\usepackage[utf8]{inputenc}
\usepackage{fontspec}
\usepackage{amsmath}
%\usepackage[spanish]{babel}

\usepackage{graphicx}

\usepackage{caption}
\usepackage{subcaption}

\usepackage{hyperref}

\setmainfont[Ligatures=TeX]{Droid Serif}

\newcommand{\Jd}{\mathbf{J}}
\newcommand{\EF}{\mathbf{E}}
\newcommand{\cond}{\boldsymbol{\sigma}}
\DeclareMathOperator{\diver}{div}
\DeclareMathOperator{\grad}{grad}
\DeclareMathOperator{\iqr}{iqr}

\title{Entropy, Symbolization and Mutual Information \\
Teorical-Numerical Quarentine Dissertation}
\author{ W. P. Karel Zapfe}

\begin{document}

\maketitle

\section{Introduction}

Electroneurophysiology is strongly based on the measurement and characterization of electrical signals given by live neurons, either in vitro or in vivo. This signal is often recorded as an array of voltage measurements, done at precise high frequencies. Along the years, the sampling and precision of these measurements have increased by several orders of magnitude. Moreover, the capacity to sample hundreds or thousands of sites simultaneously exists now, sometimes spread over macroscopic areas, such as the whole scalp of the head, or concentrated in very small areas, as in MMEAs (Massive Micro-electrode arrays). Such signals are very rich, and many efforts are done to characterize their complexity. A common approach is to estimate the Entropy of the signal, and from there to use other derived measures such as Mutual Information or Transfer Entropy to search for causality between different channels.

It turns out that the very first step is sometimes done routinely and without much thought to fundamentals, and, as we shall show here, can lead to inconsistent or plain wrong results. As we shall see, Entropy in the sense of Shannon is often wrongly used, either due to conceptual laxity or by error in the numerical algorithmic design. On the other hand, slightly different calculations of Entropy may point different "complexities" in the data. As we are most interested in causality tests, we may focus on the most consistent and robust definition of Entropy. It appears that the definition of Entropy in the sense of Shannon is a tricky thing, and Shannon himself made an erroneous extrapolation of his discrete definition that is wrong. Even though this has been pointed since 1957 by E.T. Jaynes \cite{Jaynes1957}, and adequate alternatives have been shown to exist
\cite{Kullback1951, Schroeder2004}, researchers routinely take the incorrect definition as granted, without realising that does not converge nor has the same interpretation \cite{Jaynes1968}. On the other side, the correct use of Shannon's Entropy involves
a process of symbolization of the continuous data, and this introduces further
questions. As we shall see in the following, some symbolization alter drastically
the Entropy estimation, thus forcing the researchers to take a sort of Occam's approach:
take the simplest one, even if it looks really crude, so that artifacts given
by the model do not play a large role. A lot of recent work in Neuroscience take this
path, at a very extreme level, encoding the neurophysiological signals in
only two signs, above or below a threshold. This can be done on the electric potential field, separating putative spikes from absence of spikes, or in a  more
focused manner, such as taking the sign of
the difference of the signals, as in \cite{Steuer2004}, or on
the time domain, if the inter-spike interval is larger or below a certain value,
as in \cite{Steuer2001}.

In this report we shall explore the difficulties of trying to establish the
Entropy of a continuous LFP signal, and we shall explore a symbolic approach based
on the decomposition of the signals on the CSD representation. We shall also explore
the notion of block entropy and apply it to this symbolization. Finally we shall
see if those measurements can be used to produce adequate estimations of the
Mutual Information and Transfer Entropy, so that we can explore their
usability as causality tests.

\section{Shannon Entropy}

Shannon Entropy was devised as a measure of the complexity of a message encoded in discrete symbols. This is very important because its strict definition works very well in the case that the symbolization is given, but it fails ( spectacularly one may say ) if one takes instead a possible continuum of symbols, or one imposes another encoding on the symbols. Claude Shannon himself wrongly stated that the continuous case was a limit form
of his original definition \cite{Shannon1948}, and although this was corrected
by subsequent researchers in the 50's and 60's, the error still persists in statements
that claim, as an example, that ``the entropy of the Normal distribution is equal
to the log of the standard deviation'', as for example the Wikipedia article on
Normal Distribution and the references therein \cite{WikiNormal}.

Let us remember Shannon's definition. Suppose that we have a set of discrete symbols, i.e. , a series of separated, well defined values, which could be letters, digits, or any other symbol of such a kind. For simplicity we represent them as numbers, so $n$ discrete symbols would be $\{1, 2, \ldots n \}$. We call such a set an alphabet. Suppose that we have a random process that picks one by one these symbols and gives us a string of them. Such a string would be called a "message". Now let us suppose that the process is ergodic, that means, a very long "message" would be a good representation of the probabilities of the process that picks the symbols. In such a case the process would eventually be able to produce all the possible messages that those symbols permit, and their frequency's would represent the probabilities of the random process. If we knew precisely the probabilities $P$ of every symbol $s_j$, the entropy would be:
\begin{equation}\label{eqDiffEnt}
  H=-\sum_{j=1}^n P(s_j) \log (P (s_j))
\end{equation}
The base of the logarithm is usually taken as two, in which case the Entropy has as units the bit. In that case it can be interpreted as the average number of binary option question one has to make to guess symbol correctly.  A change of base of the logarithm is simply a change of units. We shall call this definition and only this one Shannon's Entropy.
We may not have the knowledge of the $P(s_j)$ directly, so we must infer them from the messages that we sample, and by doing so we may say that we are measuring the Entropy of the messages and not of the process. This has to be taken into account if the messages are not "random enough".

Shannon, in his original paper, stated that the same idea could be done by extending it to continuous distributions:
\begin{equation}
H=-\int p(x) \log (p(x)) d x
\end{equation}
for a random continuous variable and a probability distribution $p(x)$. It turns out that the limit does not converge. One can make some algebraic tricks to make it converge, but that would affect the interpretation. The fact that this looks identical to the Thermodynamic Entropy definition gives a false sense of legitimacy. Edwin T. Jaynes \cite{Jaynes1957}  made a rigorous critique of this confusion and showed how a true relationship between Thermodynamic and Shannon's Entropy can be found. Sometimes this formula is
labeled ``Differential Entropy'', but it has a hidden assumption.
The Differential Entropy would correspond to the Kullback-Leibler divergence
with respect to a homogeneous probability, and that only makes sense if one
has a finite support!
Our way out of this incorrect usage shall be to honestly accept that real measurements
of LFP do indeed have finite support into our derivations. That actually is more sensible than taking the analytical formula without criticism and is according to the nature of real experimental data. This approach leads to the so called ``Limiting density
of Discrete Points'' \cite{Jaynes1968, Jaynes2003}, a special case of the
Kullback-Leibler Divergence \cite{Kullback1959}.

\section{Binarization of the Data and Symbolic Sequences}

Authors take at face value the equation \ref{eqDiffEnt} and try to make a numerical approximation to it. This is achieved first by making an heuristic estimate to the distribution function of continuous data. The usually approach is first to make an histogram of the data. By so doing one is actually making a symbolization of the data ( called binarization, from bins, not from binary digits ), and if one applies Shannon's formula, then one is actually calculating the entropy of the distribution of symbols from that specific binarization. One could optimistically think that such a calculation could be a reasonable approach to a limiting form of the Differential Entropy, but that is not the case. A simple example suffices.
If one has a finite set of symbols ( we shall call it an alphabet from now on), such as $\{1,2\ldots,n\}$ and a random process that picks them with equal probabilities, the Shannon Entropy either of the process or of the distribution that dictates the process is 
$$
H=-\sum \frac{1}{n} \log \frac{1}{n} = \log n
$$
A simple proof shows that this has the highest entropy for a process with $n$ symbols, taking for example the entropy of one symbol to be higher or lower than the rest, and normalizing the others so that they are still equiprobable. Then It can be shown that this probability function has lower entropy.

If one takes a continuous interval over the real numbers, say, from $a$ to $b$, and one puts a flat probability distribution function over it, the Differential Entropy will be:
$$
H=-\int \frac{1}{b-a} \log \frac{1}{b-a} d x = \log (b-a)
$$
This somehow parallels the discrete Shannon entropy for equiprobable process, but it hides a trap.

Suppose that we want to obtain the above by means of sampling a process, and that we put the data in bins of
equal length. After enough samples we may have a reasonable approach to an even distribution, but we are careful
and don't take that for granted, so we approach the integral by a numerical discrete sum. Let us suppose that we have $n$ bins, and that each one turned to be more or less equally filled with data. Say that each bin holds 1/n of the total amount of data. Then we would obtain Shannon's entropy for $n$ symbols, $H=\log n$. If we would try instead to make a Riemann sum over the expression $ p(x) \log (p(x))$ we end with another result: The bins have length $\Delta x= (b-a)/n$, and $p(x)=1/(b-a)$. 
the Integral form by the next sum:
$$
\tilde{H}=-\sum p_n(x) \log p_n(x) \Delta x \\
=-\sum \frac{1}{b-a}\log (\frac{1}{b-a}) \frac{b-a}{n} \\
= \log (b-a)
$$
The Riemann sum in this case converges trivially to the Integral form. But the numerical sampling does not! In fact, if we refine the partition of our data (and have enough data to spread them over the new bins), the Shannon Entropy begins to grow, as $\log n$. What is going on? Two things. The most overlooked is that $p(x)$ *is not* a probability: it needs to be multiplied by $dx$ to become one. The other problem is that  Shannon's entropy is the entropy of \emph{the partition itself}. The partition induces probabilities (discrete ones) and is a symbolization of the data. The Riemann Integral does not depend on the partition, as long as it converges uniformly. The Entropies do not match each other because they are measuring different things.

People familiar with Ergodic Theory shall recognize that these kind of measures, which depend on the partition, are useful, but one cannot omit stating the partition when using them. Moreover, given that a partition is a symbolization, how can we chose the right partition? We cannot approach with experimental data the Riemann sum, except on very simple examples. 
In the following we shall use some experimental data
to drive the point.

\section{ First Experimental Numerical Example }

We shall show some calculations derived from using naively the mentioned approach. Our data comes from electrophysiological experiments, recorded with the BioCAM form BrainWave, a MMEA that has 4095 recording electrodes over an area of 2.27mm by 2.27 mm. An extra
electrode is used as ground value. In the data shown below, the data sampling rate was of 7.022 KHz.

\subsection{The binarization process: different criteria}

There are different rules to select adequate bin sizes in one-dimensional data. The rules try to make the histogram "smooth enough" and without gaps. We are compelled to do so, because we assume that the phenomena that produces our data is continuous in nature, and therefore, it induces a continuous probability density function. We are looking for a histogram that captures the general shape of this pdf. If the bins are to narrow, we may end up with many empty bins, or many bins with just one data point inside it. If they are too big they may not look like the shape of the underlying distribution.

Below we compare three  methods that produce different bin sizes for the same data.
The Freedman and Diaconis rule \cite{FD1981}, takes as ruler the interquantile
range. This makes the criterion stable with respect to outliers. The exact expression
is
\begin{equation}
  l_{FD}=2 \iqr / \sqrt[3]{N},
\end{equation}
where $\iqr$ is the interquantile range and $N$ is the number of data entries.
A very similar one is Scott's rule \cite{Scott1979}:
\begin{equation}
  l_{S}=3.49 \sigma / \sqrt[3]{N},
\end{equation}
where $\sigma$ is the standard deviation. If the data behaves more or less like a
Gaussian, it gives a good bin size.
We also use the ``square root'' rule, in which we use as many data bins as the
square root of the number of data, a usual rule of thumb for obtaining histograms.
Also we set up an "heuristic" bin size, in which we do not take into account anything of the sample, but we simply pick a bin size, and set up a range. In our case the range is from $-2000 \mu V$ to $2000 \mu V$, and bin size of $4.7 \mu V$.

After obtaining histograms for each channel using each of the criteria above, we
calculate Shannon's Entropy on each channel. In the figure \ref{entrocomp} we compare
the results. 
\begin{figure}
  \includegraphics[width=0.9\textwidth]{ComparaEntropias01.png}
  \caption{Comparison of Shannon's Entropy of the same data, but
    different binning. The criteria produce qualitatively different results.}
  \label{entrocomp}
\end{figure}

\subsection{Discussion of the Entropies with different bin sizes criteria}

Inspecting the above figures, it shall be obvious that Shannon's Entropy depends highly on the partition of the range of data. The different color maps don't differ only in scale: they behave qualitatively in contrary ways. Both the Scott and the Square Root show the opposite behaviour as the Freed-Diaconis and the Heuristic fixed method. In the former higher entropy is on the noise channels, the later have higher entropy on the very active CA3 region and their neighbourhood (we discuss the count-map method at the end). Why is that and what does it mean?

Let us recant some properties of Shannon Entropy. In a finite alphabet, higher entropies imply a flatter distribution. If all the symbols have the same probability, we get the maximum entropy. Also, a distribution with more symbols tends to also have a higher entropy, unless the symbols have very different probabilities. The different criteria for obtaining the optimal bin size take into account that there may be a continuous distribution underneath it, so we are optimizing in search of it. But every channel produces a different bin size. Scott and Square root methods are very sensible to outliers, (the second one is actually dependent solely on the extrema, that is, the more outliers of them all), so they tend to produce larger bins. If the channel is active, we end with some bins that are large enough to contain all noise contribution and others that contain the activity. On the other hand,the noisy channels produce smaller bins, and the distribution of noise tend to be more homogeneous, like a truncated Gaussian. This will produce a higher entropy, as we have more bins, with a relatively flatter distribution.

In the Freed Diaconis method, we use the interquantile range as a ruler, so it is more stable in presence of outliers. That makes the bin size more consistent across channels, and therefore, the active channels tend to have a larger spread of probabilities over more or less the same alphabet, giving more Entropy to channels with more diverse signals. The heuristic criterion behaves more like this, as we use the exact same bin size and number of bins for all channels.

It turns out to be that histograms that are optimized to find a continuous distribution are inconsistent to estimate the Shannon Entropy, and this is rightly so: the Differential Entropy is \emph{not} the limit of Shannon's entropy, so that any continuous distribution can not induce a consistent discrete entropies that converge to it.

This gives us the reason to explore our true limiting partition. Our apparatus has an intrinsic discrete alphabet: the measurements are given in integer multiples of $2.014 \mu V$, so our symbols are actually those discrete values of the electric Voltage. The count-map function gives us an histogram of those discrete values. The entropy thus calculated is qualitatively similar to the Freed Diaconis Criterion, but with higher values, as it maximizes the symbols. Moreover, it also avoids misplacing measurements in the same basket, leaving other symbols empty- this is the correct alphabet for measuring the Entropy of
the signal: is consistent across channels and it minimizes our assumptions. But it shall
not generalize to more complex measures, such as Mutual Information or
Transfer Entropy, as we shall see below.

\section{Block Entropy}

Another caveat, often overlooked, is that Entropy of an array of symbols is invariant to ordering. One could in principle regard a signal as a sampling of random variables, reorder them, and obtain the same distribution, and thus, the same entropy. As an example, we could take every letter, punctuation mark and space of this paragraph, garble them into nonsense, and claim that they have the same Entropy as the original paragraph, and therefore, the same Information, which is as much nonsense as the random letters. But actually this is done implicitly in a lot of numerical estimations of Entropy.

\bibliography{./BiblioReportes01}{}
\bibliographystyle{plain}

\end{document}